---
title: "randomsearch: A brief introduction"
vignette: >
  %\VignetteIndexEntry{randomsearch: A brief introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, cache = FALSE}
library(randomsearch)
set.seed(123)
knitr::opts_chunk$set(cache = TRUE, collapse = FALSE)
```

# randomsearch

The package **randomsearch** optimizes a given [smoof function](https://cran.r-project.org/package=smoof).
Use `makeSingleObjectiveFunction()` to wrap any function into a smoof function.
It can make use of parallel resources through [prallelMap](https://cran.r-project.org/package=parallelMap).
**randomsearch** has three internal operating modes:

* *fast.parallel* is the default and used if only `max.evals` is defined as termination criterion. `max.execbudget` and `target.fun.value` have to be `NULL`.
* *slow.parallel* is used if `max.execbudget` or `target.fun.value` have been set and `parallelStart` has been called without defined `level` or with `level = randomsearch.feval`.
  In this mode `par.jobs` (defaults to `parallelGetOptions()$settings$cpus`) instances are started.
  Each instance conducts a random search until the termination criterion is met.
  The instances communicate using a shared folder (`par.dir = ~/.randomsearch/`).
* *normal* is used if no parallel mode is started and if `max.execbudget` or `target.fun.value` have been set. Note, that this introduces an overhead to *fast.parallel* even without parallelization because the termination criterion is checked after each evaluation.

## Usage

Simple usage with functions that use a vector as input:

```{r simple usage}
fun = function(x) x[1]^2 + sin(x[2])
res = randomsearch(fun, lower = c(-1, -1), upper = c(2,2), minimize = TRUE, max.evals = 30)
rs = summary(res)
rs$best.x
rs$best.y

tail(as.data.frame(res))
```

### Discrete and hierarchical search spaces

For usage with more complicated search spaces you have to define a smoof function.
This allows you to tune over discrete parameters and even hierarchical parameter sets like in this example.
Here the parameter `y` is only active if `a==a2`.

```{r usage}
fun = function(x) {
  if (x$a == "a1") {
    x$x^2
  } else if (x$a == "a2") {
    sin(x$x * x$y) + 1
  }
}
obj.fun = makeSingleObjectiveFunction(
  fn = fun,
  par.set = makeParamSet(
    makeDiscreteParam("a", values = c("a1", "a2")),
    makeNumericParam("x", lower = -3, upper = 3),
    makeNumericParam("y", lower = 0, upper = 2, requires = quote(a == a1))
  ),
  minimize = TRUE
)
res = randomsearch(obj.fun, max.evals = 30)
rs = summary(res)
rs$best.x
rs$best.y

tail(as.data.frame(res))
```

### Parallel Usage
*Note:* For Windows use `parallelStartSocket()`.

```{r parallel_usage, eval = FALSE}
fun = function(x) {
  Sys.sleep(runif(1))
  x[1]^2 + sin(x[2])
}

parallelMap::parallelStartMulticore(cpus = 2, level = "randomsearch.feval")
res = randomsearch(fun, lower = c(-1, -1), upper = c(2,2), minimize = TRUE, max.execbudget = 2, max.evals = 1000)
summary(res)
```

## Multi-objective optimization

_Note:_ `randomsearch()` detects a multi-objective optimization by the length of the `minimize` vector.

```{r multi_objective, eval = TRUE}
obj.fun = function(x) c(x[1]^2 + sin(x[2]), cos(x[1]))
res = randomsearch(obj.fun, lower = c(-1, -1), upper = c(2,2), minimize = c(TRUE, TRUE), max.evals = 30)
summary(res)
```

Of course multi-objective optimization works with smoof functions as well.
A multi-objective smoof function is constructed as follows:

```{r multi_objective_smoof}
obj.fun = makeMultiObjectiveFunction(
  fn = function(x) c(x[1]^2 + sin(x[2]), cos(x[1])),
  par.set = makeNumericParamSet(len = 2, lower = -1, upper = 2),
  minimize = c(TRUE, TRUE)
)
res = randomsearch(obj.fun, max.evals = 10)
```

## Augment random search

You can simply augment the results of a random search with additional random search results by converting the old result to a `data.frame` and use it as an input for a new random search.

```{r augment}
fun = function(x) x[1]^2 + sin(x[2]) + 1
res = randomsearch(fun, lower = c(-1, -1), upper = c(2,2), minimize = TRUE, max.evals = 10)
summary(res)
des = as.data.frame(res)
res = randomsearch(fun, lower = c(-1, -1), upper = c(2,2), minimize = TRUE, max.evals = 1000, target.fun.value = -0.4)
summary(res)
```
